
# Analysis of R version release dates

```{r setup}
library(here)
library(dplyr)
library(ggplot2)
library(fpp3)
library(ggrepel)

source(here("src", "get_latest_r_versions_data.R"))
release_dates <- readRDS(here("data", paste0(Sys.Date(), "_rversions.RDS")))
```

The `release_dates` data is sourced from the {rversions} package. 
Run the `get_latest_r_versions_data.R` script in *src/* to get the latest data. 

```{r}
release_dates
```
Which versions have had the most patch_updates?

```{r}
release_dates %>% 
  filter(date > as.Date("2000-02-07")) %>% 
  distinct(minor_version, patch_updates) %>% 
  arrange(desc(patch_updates)) %>% 
  slice(1:5)
```

# Can we guess how many more patch_updates the latest R version will have? 

At the initial time of writing this, R 4.3 only had one update. 
Upon returning to this, R 4.4.0 was just released. 

We'll build a `tsibble` to take a look at patch_updates by initial release date for each minor R version.

```{r}
major_release_ts <- 
  release_dates %>% 
  mutate(
    initial_release_week = yearweek(major_release_start),
    month = yearmonth(major_release_start),
    dev_end_week = yearweek(major_release_end)
  ) %>% 
  distinct(initial_release_week, major_version, minor_releases, dev_end_week) %>% 
  as_tsibble(index = initial_release_week)

minor_release_ts <- 
  release_dates %>% 
  mutate(
    initial_release_week = yearweek(minor_release_start),
    month = yearmonth(minor_release_start),
    dev_end_week = yearweek(minor_release_end)
  ) %>% 
  distinct(initial_release_week, minor_version, patch_updates, dev_end_week) %>% 
  as_tsibble(index = initial_release_week)
```

And on to the graph... 

Major R releases

```{r}
autoplot(major_release_ts, minor_releases) + 
  geom_point() + 
  geom_label_repel(aes(label = major_version, vjust = 0.6), direction = "y", segment.color = "blue") + 
  labs(x = NULL, y = "# of minor releases", caption = "Label positions indicate initial release date for major R version") + 
  ggtitle("How many minor releases should be expected for the latest major version of R?") + 
  scale_x_yearmonth(date_breaks = "2 years", date_label = "%Y") + 
  scale_y_continuous(limits = c(0, 15), expand = c(0.02, 0), breaks = seq(0, 15, by = 2)) + 
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = 2)
  )
```
```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_r_major_version_ts.png")),
  width = 15,
  height = 6
  )
```


Minor releases.. 

```{r fig.height=7, fig.width=12}
autoplot(minor_release_ts, patch_updates) + 
  geom_point() + 
  geom_label_repel(aes(label = minor_version, vjust = 0.6), direction = "y", segment.color = "blue") + 
  labs(x = NULL, y = "# of patch updates", caption = "Label positions indicate initial release date for minor R version") + 
  ggtitle("How many R 4.4 patch updates should be expected this year?") + 
  scale_x_yearmonth(date_breaks = "1 year", date_label = "%Y") + 
  scale_y_continuous(limits = c(0, 5.25), expand = c(0.01, 0)) + 
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = 2)
  )
```

```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_r_minor_version_ts.png")),
  width = 15,
  height = 6
  )
```

# What months have typically had the most patch_updates? 

We can look at the total number of patch_updates that have occurred in each month. 
We're currently in August 2023, so this might suggest when we can expect patch_updates to occur for the latest R version. 

```{r}
release_dates %>% 
  mutate(
    month = yearmonth(date),
    mo = lubridate::month(date, label = TRUE),
    yr = lubridate::year(date) %>% as.factor()
  ) %>% 
  group_by(mo) %>% 
  mutate(patch_updates = n()-1) %>% 
  distinct(mo, patch_updates) %>% 
  ggplot(aes(x = mo, y = patch_updates)) +
  geom_col() + 
  theme_classic() + 
  scale_y_continuous(expand = c(0,0)) + 
  labs(x = NULL, y = "total # of R releases", title = "End of year patch_updates are more likely to happen in October and December")
```

```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_minor_release_months.png")),
  width = 7,
  height = 5
  )
```


# What months have new R versions been initially released in?  

```{r}
minor_release_ts %>% 
  mutate(month = month(initial_release_week, label = TRUE)) %>% 
  count(month, sort = TRUE)
```

Interesting, so no new R versions have ever been released in January or July!








