---
title: 'Analysis of R version release dates'
output: 'github_document'
---

We'll use the `r_versions()` function from {rversions} to get past and present versions in a dataframe. 

Since R 1.4.1, version numbers consist of three numbers: major, minor and patch.

So we'll start with R 1.4.1. 

The `get_latest_r_versions_data.R` script in */src* will format names and add date variables used for plots. 

```{r setup, echo=TRUE, results="hide", message = FALSE}
library(here)
library(dplyr)
library(ggplot2)
library(rversions)
library(tsibble)
library(ggrepel)

# custom script to createand save tidy data in local sesh
source(here("src", "get_latest_r_versions_data.R"))

# load tidy data
release_dates <- readRDS(here("data", paste0(Sys.Date(), "_rversions.RDS")))
```

Take a... 

```{r}
glimpse(release_dates)
```
Which versions have had the most patch updates?

```{r}
release_dates %>% 
  distinct(minor_version, patch_updates) %>% 
  arrange(desc(patch_updates)) %>% 
  slice(1:5)
```

# Can we guess how many more patch_updates the latest R version will have? 

At the initial time of writing this, R 4.5 had `r release_dates[release_dates$minor_version=="R 4.5",]$patch_updates` updates. 

We'll use a `tsibble` to take a look at updates by initial release date for each major and then minor R version.


```{r major-versions-plot, fig.height=7, fig.width=12}
plot_majors <- 
  release_dates %>% 
  distinct(major_release_start, major_version, minor_releases) %>% 
  as_tsibble(index = major_release_start) %>% 
  ggplot(aes(x = major_release_start, y = minor_releases)) + 
  geom_line() + 
  geom_label_repel(aes(label = major_version, vjust = 0.6), direction = "y", segment.color = "blue") + 
  labs(x = NULL, y = "# of minor releases", caption = "Label positions indicate initial release date for major R version") + 
  ggtitle("How many minor releases should be expected for the latest major version of R?") + 
  scale_x_yearmonth(date_breaks = "2 years", date_label = "%Y") + 
  scale_y_continuous(limits = c(0, 12), expand = c(0.02, 0), breaks = seq(0, 12, by = 2)) + 
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = 2)
  )

plot_majors
```

And we save it...


```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_r_major_version_ts.png")),
  width = 15,
  height = 6
  )
```


Minor releases graph version

```{r minor-versions-plot, fig.height=7, fig.width=12}
plot_minors <- 
  release_dates %>% 
  distinct(minor_release_start, minor_version, patch_updates) %>% 
  as_tsibble(index = minor_release_start) %>% 
  ggplot(aes(x = minor_release_start, y = patch_updates)) + 
  geom_line() + 
  geom_label_repel(aes(label = minor_version, vjust = 0.6), direction = "y", segment.color = "blue") + 
  labs(x = NULL, y = "# of patch updates", caption = "Label positions indicate initial release date for minor R version") + 
  ggtitle("How many R 4.5 patch updates should be expected this year?") + 
  scale_x_yearmonth(date_breaks = "1 year", date_label = "%Y") + 
  scale_y_continuous(limits = c(0, 19), expand = c(0.01, 0), breaks = seq(0, 20, by = 2)) + 
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = 2)
  )

plot_minors
```

```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_r_minor_version_ts.png")),
  width = 15,
  height = 6
  )
```

# What months have typically had the most patch updates? 

We can look at the total number of patch_updates that have occurred in each month. 
We're currently in August 2023, so this might suggest when we can expect patch_updates to occur for the latest R version. 

```{r patch-update-months}
release_dates %>% 
  mutate(
    month = yearmonth(minor_release_start),
    mo = lubridate::month(minor_release_start, label = TRUE),
    yr = lubridate::year(minor_release_start) %>% as.factor()
  ) %>% 
  group_by(mo) %>% 
  mutate(patch_updates = n()-1) %>% 
  distinct(mo, patch_updates) %>% 
  ggplot(aes(x = mo, y = patch_updates)) +
  geom_col() + 
  theme_classic() + 
  scale_y_continuous(expand = c(0,0)) + 
  labs(x = NULL, y = "total # of R releases", title = "End of year patch updates are more likely to happen in October")
```

```{r}
ggsave(
  here("publish", paste0(Sys.Date(), "_minor_release_months.png")),
  width = 7,
  height = 5
  )
```


That's all for now. 







